---
title: "Diagnosing the identification of the model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Diagnosing-identification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = F
)
```

This document covers how to use the function `diagnose_identification` to diagnose the identification of the model numerically.

Basically, `diagnose_identification` takes three steps.

1. Approximate the Fisher information matrix.

2. Find the condition number and the rank of the Fisher information matrix and asymptotic variance of parameters.

3. Raise the flag that warns the large extent of the identification in a given model.

It requires the five arguments below.

1. `jags_code`

2. `data`

3. `variable.names`

4. `ll_fun`: A log-likelihood function that takes `data` and `theta` as arguments. The type and structure of `data` must match the return type of `data_fun`.

5. `data_fun`: A function that creates data with a given `theta` (a vector of parameters). `data_fun` will take an argument `args`, which includes `theta` and necessary input for generating data.


## Revisiting cross-sectional studies data example

Apply this function to the model and data in [Sampling-cross](Sampling-cross.html)

```{r setup}
library(pimbayes)
```

```{r}
ll_cross <- function(data, theta){
  se = theta[1]
  sp = theta[2]
  pi = theta[3:5]
  pi = c(pi, 1-sum(pi))
  eta =  c(se*pi[1]+(1-sp)*pi[3],
           se*pi[2]+(1-sp)*pi[4],
           sp*pi[3]+(1-se)*pi[1],
           (1-se)*pi[2]+sp*pi[4])
  log(dmultinom(data, prob = eta))
}

data_cross <- function(args) {
  theta = args$theta
  n = args$n
  se = theta[1]
  sp = theta[2]
  pi = theta[3:5]
  pi = c(pi, 1-sum(pi))
  eta =  c(se*pi[1]+(1-sp)*pi[3],
           se*pi[2]+(1-sp)*pi[4],
           sp*pi[3]+(1-se)*pi[1],
           (1-se)*pi[2]+sp*pi[4])
  as.vector(rmultinom(1, size = n, prob = eta))
}

jags_code <- "model { 
y[1:4] ~ dmulti(eta[1:4], n)
    eta[1:4] <- c(se*pi[1]+(1-sp)*pi[3],
      se*pi[2]+(1-sp)*pi[4],
      sp*pi[3]+(1-se)*pi[1],
      (1-se)*pi[2]+sp*pi[4])
      pi[1:4] ~ ddirch(c(1,1,1,1))
    se ~ dbeta(25,3)
    sp ~ dbeta(30,1.5)
  }"
data(small_cross)
data = list(y = small_cross, n = sum(small_cross))
diagnose_identification(jags_code, data,
                        c("se","sp","pi[1]","pi[2]","pi[3]"),
                        ll_cross, data_cross)
```


