---
title: "Sampling with ISTP manually"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sampling-cross-manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pimbayes)
```

If required functions and variable names could be provided, a function `istp.manual` could perform sampling with the ISTP algorithm.

Required arguments are:

1. `posterior_fun`: A function generating samples from convenience posteriors, which is proportional to the multiplication of convenience priors and likelihoods.
2. `h_inv`: $h^{-1}(\phi,\lambda) = \theta$.
3. `N`: Sample size
4. `weight_fun`: The function that calculates weight. It will take `theta_matrix`, a matrix of posterior samples of parameters in original parameterization, and `deter`, a vector of determinants.
5. `var_name`: A vector of names of variables.

Let's apply this function to the example.
Please see [Sampling-cross](Sampling-cross.html) for the example specification.

#### Inputs
```{r}
data(small_cross)
data = small_cross
posterior_fun<-function(N){
  eta<- rdirichlet(N,data+1)
  se <- rbeta(N,25,3)
  sp <- rbeta(N,30,1.5)
  phi_lambda_matrix = cbind(se,sp,eta[,1:3])
  colnames(phi_lambda_matrix) = c("se","sp", paste0("eta[",1:3,"]"))
  return(phi_lambda_matrix)
}
valid_A <-function(theta){
  pi = theta[3:5]
  pi = c(pi,1-sum(pi))
  all(pi>=0 & pi <=1)
}
weight_fun<-function(theta_matrix, deter){
  apply(theta_matrix, 1,
        function(theta) ifelse(valid_A(theta),(theta[1]+theta[2]-1)^(-2),0))
}

h_cross_inv <- function(phiLambda) {
  eta11 = phiLambda[3]
  eta12 = phiLambda[4]
  eta21 = phiLambda[5]
  se = phiLambda[1]
  sp = phiLambda[2]
  eta22 = 1 - eta11 - eta12 - eta21
  denom = se+sp-1
  pi11 = 1/denom*(sp*eta11 + (sp-1)*eta21)
  pi12 = 1/denom*(sp*eta12 + (sp-1)*eta22)
  pi21 = 1/denom*((se-1)*eta11 + se*eta21)
  return(c(se,sp,pi11,pi12,pi21))
}
var_name = c("se","sp",paste0("pi","[",c(11,12,21),"]"))
results = istp.manual(posterior_fun, h_cross_inv, 15000, weight_fun, var_name)
```

#### Results
```{r}
results$ess

head(results$samples)
```



