---
title: "ISTP sampling for incomplete multivariate binary data"
output: rmarkdown::html_vignette
bibliography: biblio.bib  
vignette: >
  %\VignetteIndexEntry{Sampling-sat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
\newcommand{\xdes}{\mathbf{X}_{des}}
\DeclareMathOperator{\logit}{logit}
\DeclareMathOperator{\expit}{expit}

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(digits = 3)
```

```{r setup}
library(pimbayes)
```

Package \texttt{pimbayes} supports the importance sampling with transparent reparameterization and convenience priors on an example of saturated logistic regression with incomplete outcome data introduced in [@lee-2024-issues].

The data have three components: the binary response variable $Y$, the binary independent variables $\mathbf{X}  = (X_1, \dots, X_p)^T$, and the indicator variable $R$, which indicates the missingness of $Y$. The binary variable $R$ is 1 when $Y$ is observed and $0$ when $Y$ is unobserved.

To ease and simplify the calculation and structure of the model, a design matrix of $\mathbf{X}$ and a vector $\xdes \in \mathbb{R}^{2^p}$ representing a row of it is used.
In mathematical notation, $\xdes  = (X_{0\dots0},\dots,X_{1,\dots1})$ and each entry $X_{i_{1},i_{2},\dots ,i_{p}} = 1$ is 1 if and only if $X_{1} = i_1, X_2 = i_{2}, \dots, X_p = i_p$.


\newcommand{\bbR}{\mathbb{R}}

There are multiple vectors of parameters of interest,
- $\alpha  = (p_{0\dots0}, \dots,p_{1\dots1})^T$  where $\alpha$ belongs to the probability simplex on $2^p$ categories and  
$p_{i_{1},i_{2},\dots, i_{p}}  = Pr(X_{1} = i_1, X_2 = i_{2}, \dots, X_p = i_p)$;
- $\beta  \in \mathbb{R}^{2^p};$
- $\gamma \in \mathbb{R}^{2^p};$
- $\delta \in \mathbb{R}^{2^p}.$

Finally, the model is specified with

$\mathbf{X}_{des}  \sim Multinom(\alpha,1)$
$(Y \mid \mathbf{X}_{des})  \sim Bern(\expit(\mathbf{X}_{des} {\beta}));\\
(R \mid \mathbf{X}_{des}, Y = 0)
 \sim Bern(\expit(\mathbf{X}_{des} \gamma));\\
(R \mid \mathbf{X}_{des}, Y = 1)
 \sim Bern(\expit(\mathbf{X}_{des} (\expit(\logit(\gamma+\delta))))).$
 
Priors for these parameters are
\begin{equation}
\begin{aligned}
        \alpha & \sim Dir(\mathbf{1}_{2^p}), &  \\
        \beta_{i} & \sim Unif(0,1) & (i = 0, \dots ,2^p-1), \\
        \gamma_{i} & \sim Unif(0,1) & (i = 0, \dots ,2^p-1), \\
    \delta_{i} & \sim Normal(0, \sigma^2) &  (i = 0, \dots ,2^p-1). 
\end{aligned}
\end{equation}


## Data

Two mock datasets are available, \texttt{small\_sat} and \texttt{sat}. The codes for creating the mock data are also available in the folder \texttt{data-raw}.
Data $\texttt{small\_sat}$ is created from a small simulation with $n =100$ and $p =2$.

```{r}
data(small_sat)
head(small_sat$x)
head(small_sat$y)
```

## JAGS Code

```{r}
jags_code  <- "
model{
for (i in 1:n) {
    x[i, 1:p2] ~ dmulti(px[1:p2],1)
    y[i] ~ dbern(inprod(x[i,1:p2], py_x[1:p2]))
    r[i] ~ dbern(pr[i])
    pr[i] <- (y[i]) * inprod(x[i,1:p2], pr_xy1[1:p2]) + abs(y[i]-1) * inprod(x[i,1:p2], pr_xy0[1:p2])
}
for(i in 1:p2){
    pr_xy1[i] <- ilogit(log_or_r[i] + logit(pr_xy0[i]))
}
  px[1:p2] ~ ddirich(Int64wrapper(ones(p2)))
  for(i in 1:p2) {
      py_x[i] ~ dunif(0, 1)
      pr_xy0[i] ~ dunif(0, 1)
      log_or_r[i] ~ dnorm(0, tau)
    }
}"
```


## Sampling

When data and JAGS code are given with the desired posterior sample size, the function \texttt{sampling()} will return posterior samples along with ESS.
```{r message = F}
n = length(small_sat$r)
p = log2(dim(small_sat$x)[2])
sigma = 0.5 # Hyperparameter
data <- c(small_sat, list(n = n, p2 = 2^p, tau = 1/sigma^2))
results = sampling(jags_code, data, type = "sat", N = 1000)
results$ess
head(results$samples[,1:7],3)
head(results$samples[,8:11],3)
head(results$samples[,12:15],3)
```



# References


