---
title: "Sampling with Pigeons.jl"
output: rmarkdown::html_vignette
bibliography: biblio.bib  
vignette: >
  %\VignetteIndexEntry{Sampling-pigeons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r setup}
library(pimbayes)
```

## Sampling with Pigeons.jl

It is possible to obtain MCMC samples from posterior distributions using the `julia` package `Pigeons.jl` [@surjanovic2023pigeons], which may have better MCMC performance with the partially identified model thanks to the non-reversible PT [@syed-2021-Non-Reversible].

### Multithreaded option for `pigeons` sampling

Unlike other functions in this `pimbayes` package, the number of threads for the `julia` session is crucial due to the `multithreaded` options. The default option of `JuliaCall` [@R-JuliaCall], which bridges between `julia` and `R`, is a single-thread session. Since it is impossible to change the number of threads in the `julia` session once it starts, a restart of R is required if the `multithreaded` option is required but has already been initialized without setting the number of threads. 

When R starts or restarts, the user can set up a fresh `julia` session with a given number of threads. 

```{r}
init_julia_session_multi(threads = 4)
```

Next, the user can use `sampling()` with the desired number of threads. We use the small number for `n_rounds` to reduce the rendering times of this example code. In practice, using a larger number of `n_rounds` is ideal to ensure a sufficient number of iterations, $2^n$.

```{r}
data(small_count)
sigma = 0.5 #Hyperparameter, prior variance of alpha1
data = c(small_count, list(n = length(small_count$r), tau = 1/sigma^2))
jags_code <- "model{
  for (i in 1:n) {
    r[i] ~ dbern(pr[i])
    pr[i] <- ilogit(y[i] * alpha1 + alpha0)
    y[i] ~ dpois(mu)
  }
    mu ~ dgamma(1,1)
    alpha0 ~ dnorm(0, 0.1)
    alpha1 ~ dnorm(0, tau)
}
"
pt = sampling(jags_code, 
    data, 
    "pigeons", extract_variable_names = c("mu","alpha0","alpha1"), 
    n_chains =  4,
    n_rounds = 4)   
```


Using functions `pigeons_samples` and `pigeons_plots`, it is possible to check the results. The following matrix shows the first six rows of the output, which are samples from the posterior distributions. 

```{r}
head(pigeons_samples(pt))
```

The software also supports trace and density plots.

```{r}
pigeons_plots(pt)
```


There are two options for the plot. The default option of `Pigeons.jl` is saving the samples of target distribution only as above. However, it is also possible to save all chains with `extended_traces` option. When the option is enabled, the plot shows each tempered chain in different colours as below.

```{r}
pt = sampling(jags_code, data, "pigeons", extract_variable_names = c("mu","alpha0","alpha1"), n_chains =  4, n_rounds = 4, extended_traces = T)
pigeons_plots(pt)
```

## Reference


